---
title: "KPMP scRNAseq Data Analysis for Soumya Maity"
subtitle: "Exploration of MTAP and APRT Expression"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: false
    number_sections: no
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=T, error=FALSE, message=FALSE, warning=FALSE}

if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("SeuratDisk")) {install.packages("SeuratDisk"); require("SeuratDisk")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("paletteer")) {install.packages("paletteer"); require("paletteer")} # color palette
if (!require("grDevices")) {install.packages("grDevices"); require("grDevices")} # for grDevices palette
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("tibble")) {install.packages("tibble"); require("tibble")} # for table transformation
if (!require("geneName")) {install.packages("geneName"); require("geneName")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("ggupset")) {install.packages("ggupset"); require("ggupset")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("viridis")) {install.packages("viridis"); require("viridis")}

library(SeuratData)
library(openxlsx)
library(gplots)
library(ggvenn)

```

# Introduction

This file was made to explore the expression of MTAP and APRT in the KPMP dataset in Healthy vs. AKI/CKD as well as after subsetting PT cells and classifying based on expression of KIM1 (HAVCR1).

## Additions to Analysis of single-cell data on 11.16.23

# Notes

In the 2024 KPMP scRNAseq (and I think also in the snRNAseq file) there is some mismatch between the *subclass.l1* and *subclass.l2* annotations. For example, subsetting subclass.l1 == "PT" and then looking at subclass.l2, there are some cells that are not PT.

Working with the KPMP scRNAseq file: **PREMIERE_Alldatasets_08132021.h5Seurat**

# Interpretation from JWN (11.13.24)

1) MTAP and APRT are weakly expressed in this dataset (Especially true for MTAP) (**Section 4**).
2) MTAP and APRT appear more abundant in PT cells experiencing AKI (but the sampling of cells is pretty small) (**Section 5**).
3) MTAP *BUT NOT* APRT is more abundant in PT cells that express KIM1 (indicating injury) (**Section 6**).

# 2021 scRNASeq KPMP Object

## Load KPMP object (2021 data file)

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Sys.time()
KPMP <- LoadH5Seurat(here("PREMIERE_Alldatasets_08132021.h5Seurat")) 
Sys.time()

head(KPMP@meta.data)

DimPlot(KPMP, group.by = "subclass.l1")

KPMP

```

## Add Clinical information into meta.data

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

meta <- read.csv(here("OpenAccessClinicalData.csv"))

df <- FetchData(KPMP, "SpecimenID") %>% rownames_to_column(var = "CellID")

df <- left_join(df, meta, by = c("SpecimenID" = "Participant.ID")) %>% column_to_rownames(var = "CellID")

KPMP <- AddMetaData(KPMP, df)

head(KPMP@meta.data)

unique(KPMP@meta.data$Enrollment.Category)

KPMP@meta.data$Enrollment.Category <- factor(KPMP@meta.data$Enrollment.Category, levels = c("Healthy Reference", "AKI", "CKD"))

```


# Overall Expression of MTAP and APRT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Idents(KPMP) <- KPMP@meta.data$subclass.l1

FeaturePlot(KPMP, "MTAP", order = TRUE)

VlnPlot(KPMP, features = "MTAP", group.by = "subclass.l1") + ggtitle("MTAP Expression in all cells")

VlnPlot(KPMP, features = "MTAP", group.by = "subclass.l1") + ggtitle("MTAP Expression in all cells")


FeaturePlot(KPMP, "APRT", order = TRUE)

VlnPlot(KPMP, features = "APRT", group.by = "subclass.l1") + ggtitle("APRT Expression in all cells")

```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10}

DotPlot(
  KPMP,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "subclass.l1",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("MTAP Expression in all cells")

DotPlot(
  KPMP,
  assay = NULL,
  features = c("APRT"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "subclass.l1",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("APRT Expression in all cells")

# rotate x-axis labels on the plot above





  

```


# Expression of MTAP and APRT in AKI

## MTAP

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(KPMP, features = "MTAP", group.by = "Enrollment.Category", pt.size = 0.1) + ggtitle("MTAP in all cells")

VlnPlot(KPMP, features = "MTAP", idents = "PT", group.by = "Enrollment.Category", pt.size = 0.1) + ggtitle("MTAP in PT cells")


DotPlot(
  KPMP,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "Enrollment.Category",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("MTAP in all cells")

DotPlot(
  KPMP,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = "PT",
  group.by = "Enrollment.Category",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("MTAP in PT cells")




```

## APRT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(KPMP, features = "APRT", group.by = "Enrollment.Category", pt.size = 0.1) + ggtitle("APRT in all cells")

VlnPlot(KPMP, features = "APRT", idents = "PT", group.by = "Enrollment.Category", pt.size = 0.1) + ggtitle("APRT in PT cells")


DotPlot(
  KPMP,
  assay = NULL,
  features = c("APRT"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "Enrollment.Category",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("APRT in all cells")

DotPlot(
  KPMP,
  assay = NULL,
  features = c("APRT"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = "PT",
  group.by = "Enrollment.Category",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("APRT in PT cells")



```


# Clasify Cells by Kim1 Expression

## Subset PT cells

Kim1 = HAVCR1

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

KPMP_PT <- subset(KPMP, idents = c("PT"))

DimPlot(KPMP_PT, group.by = "subclass.l2") + ggtitle("KPMP PT cells")

VlnPlot(KPMP_PT, features = "HAVCR1", group.by = "subclass.l2", pt.size = 0.1) + ggtitle("HAVCR1 in PT cells")

VlnPlot(KPMP_PT, features = "HAVCR1", group.by = "subclass.l2", split.by = "Enrollment.Category", pt.size = 0.1) + ggtitle("HAVCR1 in PT cells")

KPMP_PT@meta.data$KIM1status <- ifelse(GetAssayData(KPMP_PT, assay = "RNA", slot = "data")["HAVCR1", ] > 0, "yes", "no")

DimPlot(KPMP_PT, group.by = "KIM1status", label = T)

table(KPMP_PT@meta.data$KIM1status, KPMP_PT@meta.data$subclass.l2)

table(KPMP_PT@meta.data$KIM1status, KPMP_PT@meta.data$Enrollment.Category)

```

### MTAP

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(KPMP_PT, features = "MTAP", group.by = "KIM1status", pt.size = 0.1) + ggtitle("MTAP in PT cells by KIM1 Expression")

DotPlot(
  KPMP_PT,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = "PT",
  group.by = "KIM1status",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("MTAP in PT cells by KIM1 Expression")



```

### APRT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(KPMP_PT, features = "APRT", group.by = "KIM1status", pt.size = 0.1) + ggtitle("APRT in PT cells by KIM1 Expression")

DotPlot(
  KPMP_PT,
  assay = NULL,
  features = c("APRT"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = "PT",
  group.by = "KIM1status",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("APRT in PT cells by KIM1 Expression")


```

## Only in AKI PT Cells

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Idents(KPMP_PT) <- KPMP_PT@meta.data$Enrollment.Category

KPMP_PT2 <- subset(KPMP_PT, idents = c("AKI"))

DimPlot(KPMP_PT2, group.by = "subclass.l2") + ggtitle("KPMP AKI PT cells")

VlnPlot(KPMP_PT2, features = "HAVCR1", group.by = "subclass.l2", pt.size = 0.1) + ggtitle("HAVCR1 in AKI PT cells")

DimPlot(KPMP_PT2, group.by = "KIM1status", label = T)

table(KPMP_PT2@meta.data$KIM1status, KPMP_PT2@meta.data$subclass.l2)

table(KPMP_PT2@meta.data$KIM1status, KPMP_PT2@meta.data$Enrollment.Category)

```


### MTAP

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(KPMP_PT2, features = "MTAP", group.by = "KIM1status", pt.size = 0.1) + ggtitle("MTAP in AKI PT cells by KIM1 Expression")

DotPlot(
  KPMP_PT2,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "KIM1status",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("MTAP in AKI PT cells by KIM1 Expression")



```

### APRT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

VlnPlot(KPMP_PT2, features = "APRT", group.by = "KIM1status", pt.size = 0.1) + ggtitle("APRT in AKI PT cells by KIM1 Expression")

DotPlot(
  KPMP_PT2,
  assay = NULL,
  features = c("APRT"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "KIM1status",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + ggtitle("APRT in AKI PT cells by KIM1 Expression")


```

# Figures for Manuscript

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

head(KPMP@meta.data)

```

### Pre-Figure Work

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

levels(KPMP@meta.data$subclass.l1)

KPMP@meta.data$subclass.l1 <- factor(KPMP@meta.data$subclass.l1, levels = c("PT", "DTL", "ATL/TAL", "TAL", "DCT", "CNT", "PC", "IC", "POD", "PEC", "EC", "Interstitial", "Immune"))

DimPlot(KPMP, group.by = "subclass.l1") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))


KPMP_subset <- subset(KPMP, idents = c("PT"))

DimPlot(KPMP_subset, group.by = "subclass.l1") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20))

Idents(KPMP_subset) <- KPMP_subset@meta.data$Enrollment.Category

KPMP_subset <- subset(x = KPMP_subset, idents = c("Healthy Reference", "AKI"))

DimPlot(KPMP_subset, group.by = "subclass.l1", split.by = "Enrollment.Category") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20))


DimPlot(KPMP_subset, group.by = "subclass.l2") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20))

KPMP_subset@meta.data$subclass.l2 <- factor(KPMP_subset@meta.data$subclass.l2, levels = c("PT-S1/S2", "PT-S3", "aPT", "dPT", "dPT/DTL", "cycEPI"))

DimPlot(KPMP_subset, group.by = "subclass.l2") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))






```
## Figure 1

### Figure 1A

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DimPlot(KPMP, group.by = "subclass.l1") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

DimPlot(KPMP_subset, group.by = "subclass.l2") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

```

### Figure 1b

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10, fig.height=6}

VlnPlot(
  object = KPMP_subset, 
  features = "MTAP", 
  group.by = "subclass.l2", 
  split.by = "Enrollment.Category"
) +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ylab("MTAP Expression") +
  ggtitle("MTAP Expression Across PT Cells")


```

### Figure 1c

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10, fig.height=6}

VlnPlot(
  object = KPMP_subset, 
  features = "HAVCR1", 
  group.by = "subclass.l2", 
  split.by = "Enrollment.Category"
) +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ylab("HAVCR1 Expression") +
  ggtitle("HAVCR1 Expression Across PT Cells")


```

### Figure 1d

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10, fig.height=4}

DotPlot(
  KPMP_subset,
  assay = NULL,
  features = c("MTAP", "HAVCR1"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  group.by = "Enrollment.Category",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ggtitle("HAVCR1 and MTAP Expression") +
  xlab(label = "")

```

### Figure 1d

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

KPMP_subset@meta.data$KIM1status <- ifelse(GetAssayData(KPMP_subset, assay = "RNA", slot = "data")["HAVCR1", ] > 0, "yes", "no")

DimPlot(KPMP_subset, group.by = "KIM1status", order = T) + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("PT Cells Categorized by KIM1 Expression")

FeaturePlot(KPMP_subset, "HAVCR1", order = T) + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        plot.title = element_text(hjust = 0.5))+
  ggtitle("KIM1 Expression")


```

### Figure 1e

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10, fig.height=5}

DotPlot(
  KPMP_subset,
  assay = "RNA",  # Specify the assay explicitly if applicable
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  group.by = "KIM1status",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = TRUE,
  scale.by = "radius"
) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_text(size = 20, colour = "black"),
    axis.title.y = element_text(size = 15, colour = "black"), # Force y-axis title
    axis.text.x = element_text(size = 25, colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ggtitle("MTAP Expression Based on KIM1 Expression") +
  ylab("Does Cell Express KIM1?") +  
  xlab("")


```

## Figure S1

### Figure S1a

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

KPMP@meta.data$subclass.l1 <- factor(KPMP@meta.data$subclass.l1, levels = c("PT", "DTL", "ATL/TAL", "TAL", "DCT", "CNT", "PC", "IC", "POD", "PEC", "EC", "Interstitial", "Immune"))

Idents(KPMP) <- KPMP@meta.data$Enrollment.Category

KPMP_subset2 <- subset(x = KPMP, idents = c("Healthy Reference", "AKI"))

DimPlot(KPMP_subset2, group.by = "subclass.l1") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        plot.title = element_text(hjust = 0.5)) +
    ggtitle("Healthy Reference and AKI Cells")


```

### Figure S1b

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=4, fig.width=8}

DotPlot(
  KPMP_subset2,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  group.by = "Enrollment.Category",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ggtitle("MTAP Expression by Disease") +
  xlab(label = "")

```

### Figure S1c

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=12, fig.height=4}

KPMP_subset3 <- subset(x = KPMP, idents = c("Healthy Reference"))

KPMP_subset4 <- subset(x = KPMP, idents = c("AKI"))

DotPlot(
  KPMP_subset3,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  group.by = "subclass.l1",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ggtitle("MTAP Expression by Cell Type in Healthy Reference") +
  xlab(label = "")

DotPlot(
  KPMP_subset4,
  assay = NULL,
  features = c("MTAP"),
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  group.by = "subclass.l1",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, colour = "black"),  # X-axis font color
    axis.text.y = element_text(colour = "black")   # Y-axis font color
  ) +
  ggtitle("MTAP Expression by Cell Type in AKI") +
  xlab(label = "")


```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

head(KPMP_subset@meta.data)

table(KPMP_subset@meta.data$SpecimenID, KPMP_subset@meta.data$Enrollment.Category)

# Create the contingency table
enrollment_table <- table(KPMP_subset@meta.data$SpecimenID, KPMP_subset@meta.data$Enrollment.Category)

# Count the number of unique samples (rows with non-zero entries) for each category
sample_counts <- colSums(enrollment_table > 0)
print(sample_counts)

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Sys.time()

sessionInfo()

```
